---
title: "Bagging and Random Forest"
author: "Fortuna Zhang"
output: pdf_document
---

## Bagging and Random Forest
Variables: 
There are 14 attributes in each case of the dataset. They are:
CRIM - per capita crime rate by town
ZN - proportion of residential land zoned for lots over 25,000 sq.ft.
INDUS - proportion of non-retail business acres per town.
CHAS - Charles River dummy variable (1 if tract bounds river; 0 otherwise)
NOX - nitric oxides concentration (parts per 10 million)
RM - average number of rooms per dwelling
AGE - proportion of owner-occupied units built prior to 1940
DIS - weighted distances to five Boston employment centres
RAD - index of accessibility to radial highways
TAX - full-value property-tax rate per $10,000
PTRATIO - pupil-teacher ratio by town
B - 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town
LSTAT - % lower status of the population
MEDV - Median value of owner-occupied homes in $1000's

## Training and Test set
```{r}
library(MASS)

set.seed(1)

names(Boston)

train = sample(1:nrow(Boston),nrow(Boston)/2)

boston.test=Boston[-train,"medv"]
```

## Bagged tree
```{r}
# random forest 
library(randomForest)

set.seed(1)
# Bagged tree
# mtry: # of predictors used in branching; importance: variable importance score
bag.boston = randomForest(medv~., data=Boston, subset=train, mtry=13, importance=TRUE)

bag.boston
```

## Visualization of bagged tree
```{r}
# prediction of bagged tree
yhat.bag = predict(bag.boston, newdata = Boston[-train,])
# Visualization
plot(yhat.bag,boston.test)
# test MSE
mean((yhat.bag-boston.test)^2)
```

## Bagged tree with 25 trees
```{r}
# Build bagged tree again, ntree: # of trees
bag.boston = randomForest(medv~.,data = Boston, subset=train, mtry=13, ntree=25)

yhat.bag = predict(bag.boston,newdata = Boston[-train,])

mean((yhat.bag-boston.test)^2)
```
Test MSE is worse since 25 trees are not sufficient. 

## Random forest
```{r}
# choose candidate predictors randomly for every branching
set.seed(1)
# mtry=6: choose 6 predictors for branching
rf.boston = randomForest(medv~.,data=Boston,subset=train,mtry=6,importance = TRUE)

yhat.rf = predict(rf.boston,newdata = Boston[-train,])

mean((yhat.rf-boston.test)^2)
```
Random forest with less predictors is better. Test MSE lower than bagged tree.

## Visualization of random forest
```{r}
# variable importance plot
varImpPlot(rf.boston)
# variable importance score
importance(rf.boston)
```
%IncMSE: how much MSE will worse if take out that predictor
IncNodePurity: how much impure the node is if take out that predictor
In this case, rm, lstat are important predictors. 

## Boosted tree
```{r}
# Boosted tree
library(gbm)

set.seed(1)
# target variable is numeric: "gaussian"; n.tress: # of trees; interaction.depth: max # of leaf nodes in each tree
boost.boston = gbm(medv~.,Boston[train,],distribution = "gaussian", n.trees=5000,interaction.depth=4)
summary(boost.boston)
```


# Visualization of boosted tree
```{r}

par(mfrow=c(1,2))
# Partial Dependency Plots
plot(boost.boston,i="rm")

plot(boost.boston,i="lstat")

par(mfrow=c(1,1))
```
Plot shows relationships between rm/lstat with target variable.

## prediction of boosted tree
```{r}
yhat.boost=predict(boost.boston,newdata=Boston[-train,],n.trees=5000)

boston.test=Boston[-train,"medv"]

mean((yhat.boost-boston.test)^2)
```

## Tune boosted using shrinkage rate
```{r}
set.seed(1)
# shrinkage***: learning rate; verbose=F: not report progress while building model

boost.boston = gbm(medv~.,data=Boston[train,],distribution="gaussian",n.trees=5000,interaction.depth=4,shrinkage=0.2,verbose=F)
# use 5000 trees 
yhat.boost=predict(boost.boston,newdata=Boston[-train,],n.trees=5000)

mean((yhat.boost-boston.test)^2)
# use 500 trees 
yhat.boost=predict(boost.boston,newdata=Boston[-train,],n.trees=500)

mean((yhat.boost-boston.test)^2)
```

500 and 5000 trees are similar. Tune shrinkage rate could improve accuracy. 
Boosted tree is the best among random forest and bagged tree.
